---
title: "Working with Charts"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working with Charts}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup}
library(rpptx)
```

```{r helper, include=FALSE}
blank_slide <- function() {
  prs    <- pptx_presentation()
  layout <- prs$slide_layouts[[6]]
  list(prs = prs, slide = prs$slides$add_slide(layout))
}
```

## Overview

Charts in rpptx embed a real Excel workbook inside the `.pptx` file — exactly as PowerPoint
does. The workflow is:

1. Create a `CategoryChartData` (or `XyChartData` / `BubbleChartData`) object
2. Add categories and series
3. Call `slide$shapes$add_chart()` with an `XL_CHART_TYPE` constant

## Category charts (bar, column, line, pie, …)

### Clustered column chart

```{r}
r <- blank_slide()
slide <- r$slide

cd <- CategoryChartData$new()
cd$categories <- c("Q1", "Q2", "Q3", "Q4")
cd$add_series("Revenue",  c(120, 145, 132, 178))
cd$add_series("Expenses", c(95,  110, 105, 130))

chart_shape <- slide$shapes$add_chart(
  XL_CHART_TYPE$COLUMN_CLUSTERED,
  left   = Inches(0.5),
  top    = Inches(0.5),
  width  = Inches(9),
  height = Inches(6),
  chart_data = cd
)
chart_shape$shape_type == MSO_SHAPE_TYPE$CHART
```

### Bar chart (horizontal)

```{r}
r <- blank_slide()
slide <- r$slide

cd <- CategoryChartData$new()
cd$categories <- c("Product A", "Product B", "Product C")
cd$add_series("Units sold", c(450, 320, 580))

slide$shapes$add_chart(
  XL_CHART_TYPE$BAR_CLUSTERED,
  Inches(0.5), Inches(0.5), Inches(9), Inches(6), cd
)
```

### Line chart

```{r}
r <- blank_slide()
slide <- r$slide

cd <- CategoryChartData$new()
cd$categories <- month.abb
cd$add_series("2023", c(40, 45, 55, 60, 70, 75, 80, 78, 65, 55, 48, 42))
cd$add_series("2024", c(42, 48, 58, 65, 72, 80, 85, 82, 70, 60, 52, 45))

slide$shapes$add_chart(
  XL_CHART_TYPE$LINE,
  Inches(0.5), Inches(0.5), Inches(9), Inches(6), cd
)
```

### Pie chart

```{r}
r <- blank_slide()
slide <- r$slide

cd <- CategoryChartData$new()
cd$categories <- c("North", "South", "East", "West")
cd$add_series("Market share", c(35, 25, 20, 20))

slide$shapes$add_chart(
  XL_CHART_TYPE$PIE,
  Inches(1), Inches(0.5), Inches(8), Inches(6), cd
)
```

### Area chart

```{r}
r <- blank_slide()
slide <- r$slide

cd <- CategoryChartData$new()
cd$categories <- paste0("Week ", 1:8)
cd$add_series("Team A", c(10, 15, 13, 20, 18, 22, 25, 24))
cd$add_series("Team B", c(8,  12, 10, 16, 14, 18, 20, 19))

slide$shapes$add_chart(
  XL_CHART_TYPE$AREA,
  Inches(0.5), Inches(0.5), Inches(9), Inches(6), cd
)
```

## Available chart types

Some commonly used `XL_CHART_TYPE` constants:

| Constant | Description |
|---|---|
| `COLUMN_CLUSTERED` | Clustered column (vertical bars) |
| `COLUMN_STACKED` | Stacked column |
| `COLUMN_STACKED_100` | 100% stacked column |
| `BAR_CLUSTERED` | Clustered bar (horizontal) |
| `BAR_STACKED` | Stacked bar |
| `LINE` | Line chart |
| `LINE_MARKERS` | Line with markers |
| `LINE_STACKED` | Stacked line |
| `PIE` | Pie chart |
| `PIE_EXPLODED` | Exploded pie |
| `AREA` | Area chart |
| `AREA_STACKED` | Stacked area |
| `XY_SCATTER` | Scatter (XY) |
| `BUBBLE` | Bubble chart |
| `DOUGHNUT` | Doughnut |
| `RADAR` | Radar / spider |

## XY (scatter) charts

XY charts need `XyChartData` which accepts (x, y) pairs per series:

```{r}
r <- blank_slide()
slide <- r$slide

cd <- XyChartData$new()
cd$add_series("Series 1",
  x_values = c(1, 2, 3, 4, 5),
  y_values = c(2.1, 3.8, 5.2, 7.1, 9.0)
)
cd$add_series("Series 2",
  x_values = c(1, 2, 3, 4, 5),
  y_values = c(1.5, 2.9, 4.1, 5.8, 7.3)
)

slide$shapes$add_chart(
  XL_CHART_TYPE$XY_SCATTER,
  Inches(0.5), Inches(0.5), Inches(9), Inches(6), cd
)
```

## Accessing chart properties

After adding a chart, retrieve the `Chart` object to adjust its properties:

```{r}
r <- blank_slide()
slide <- r$slide

cd <- CategoryChartData$new()
cd$categories <- c("A", "B", "C")
cd$add_series("S1", c(1, 2, 3))

chart_shape <- slide$shapes$add_chart(
  XL_CHART_TYPE$COLUMN_CLUSTERED,
  Inches(0.5), Inches(0.5), Inches(9), Inches(6), cd
)

chart <- chart_shape$chart
class(chart)   # "Chart"
```

### Chart title

```{r}
chart$has_title <- TRUE
chart$chart_title$text_frame$text <- "My Chart Title"
```

### Legend

```{r}
chart$has_legend <- TRUE
chart$legend$position <- XL_LEGEND_POSITION$BOTTOM
chart$legend$include_in_layout <- FALSE
```

Legend positions: `TOP`, `BOTTOM`, `LEFT`, `RIGHT`, `CORNER`.

### Axes

```{r}
cat_axis <- chart$category_axis
val_axis  <- chart$value_axis

# Axis visibility
cat_axis$visible <- TRUE
val_axis$visible <- TRUE

# Tick label position
cat_axis$tick_label_position <- "low"
val_axis$tick_label_position <- "nextTo"

# Scale
val_axis$minimum_scale <- 0
val_axis$maximum_scale <- 200
```

### Data labels

```{r}
plot <- chart$plots[[1]]
plot$has_data_labels <- TRUE
dls <- plot$data_labels
dls$show_value      <- TRUE
dls$show_category_name <- FALSE
dls$show_series_name   <- FALSE
```

## Chart type reference

```{r}
# All available chart types
length(XL_CHART_TYPE)   # 73+
```

```{r eval=FALSE}
# Common types
XL_CHART_TYPE$COLUMN_CLUSTERED
XL_CHART_TYPE$BAR_CLUSTERED
XL_CHART_TYPE$LINE
XL_CHART_TYPE$PIE
XL_CHART_TYPE$AREA
XL_CHART_TYPE$XY_SCATTER
XL_CHART_TYPE$BUBBLE
XL_CHART_TYPE$DOUGHNUT
XL_CHART_TYPE$RADAR
```

## Full example

```{r}
prs    <- pptx_presentation()
layout <- prs$slide_layouts[[6]]
slide  <- prs$slides$add_slide(layout)

# Build chart data
cd <- CategoryChartData$new()
cd$categories <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun")
cd$add_series("Actual",   c(120, 135, 128, 155, 162, 175))
cd$add_series("Forecast", c(115, 130, 140, 150, 165, 180))

# Add chart
cs <- slide$shapes$add_chart(
  XL_CHART_TYPE$LINE_MARKERS,
  Inches(0.5), Inches(0.5),
  Inches(9),   Inches(6.5),
  cd
)

# Decorate
ch <- cs$chart
ch$has_title <- TRUE
ch$chart_title$text_frame$text <- "H1 Performance"
ch$has_legend <- TRUE
ch$legend$position <- XL_LEGEND_POSITION$BOTTOM

tmp <- tempfile(fileext = ".pptx")
prs$save(tmp)
```
