---
title: "Getting Started with rpptx"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with rpptx}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(rpptx)
```

`rpptx` is an R port of the Python `python-pptx` library. It lets you create, read, and update
PowerPoint (`.pptx`) files entirely from R, with an R6-based object model that closely mirrors
the Python API.

## Creating a presentation

`pptx_presentation()` opens an existing file or — with no argument — creates a new blank presentation
based on the built-in default template.

```{r}
prs <- pptx_presentation()
```

Check the slide dimensions (in EMU — English Metric Units, 914400 per inch):

```{r}
prs$slide_width   # 10 inches
prs$slide_height  # 7.5 inches
```

Use `Inches()`, `Cm()`, or `Pt()` to convert from convenient units to EMU:

```{r}
Inches(1)     # 914400L
Cm(2.54)      # 914400L
Pt(72)        # 914400L
```

## Adding slides

Slides are added through the `Slides` collection using a slide layout as a template.
Every new blank presentation has one slide master with eleven layouts.

```{r}
# List available layout names
sapply(seq_len(length(prs$slide_layouts)), function(i) {
  prs$slide_layouts[[i]]$name
})
```

Add a blank slide (layout index 6 is "Blank" in the default template):

```{r}
layout <- prs$slide_layouts[[6]]
slide  <- prs$slides$add_slide(layout)
```

Access individual slides by 1-based index:

```{r}
length(prs$slides)    # 1
prs$slides[[1]]       # the slide we just added
```

## Adding a text box

```{r}
slide$shapes$add_textbox(
  left   = Inches(1),
  top    = Inches(1),
  width  = Inches(5),
  height = Inches(1)
)
shape <- slide$shapes[[1]]
shape$text_frame$text <- "Hello, rpptx!"
```

## Adding a rectangle

```{r}
rect <- slide$shapes$add_shape(
  MSO_AUTO_SHAPE_TYPE$ROUNDED_RECTANGLE,
  left   = Inches(1),
  top    = Inches(2.5),
  width  = Inches(3),
  height = Inches(1.5)
)
rect$text <- "A rounded rectangle"
rect$fill$solid()
rect$fill$fore_color$rgb <- RGBColor(0x46L, 0x72L, 0xC4L)
```

## Adding a picture

```{r eval=FALSE}
slide$shapes$add_picture(
  "path/to/photo.jpg",
  left   = Inches(5),
  top    = Inches(2),
  width  = Inches(3),
  height = Inches(2)
)
```

## Saving

```{r eval=FALSE}
prs$save("my_presentation.pptx")
```

Or save to a temp file:

```{r}
tmp <- tempfile(fileext = ".pptx")
prs$save(tmp)
file.exists(tmp)
```

## Opening an existing file

```{r eval=FALSE}
prs2 <- pptx_presentation("existing.pptx")
slide1 <- prs2$slides[[1]]
for (shape in slide1$shapes$to_list()) {
  cat(shape$name, "\n")
}
```

## Slide operations

### Deleting a slide

```{r}
prs2 <- pptx_presentation()
layout <- prs2$slide_layouts[[6]]
s1 <- prs2$slides$add_slide(layout)
s2 <- prs2$slides$add_slide(layout)
length(prs2$slides)   # 2
prs2$slides$delete(s1)
length(prs2$slides)   # 1
```

### Reordering slides

```{r}
prs3 <- pptx_presentation()
layout <- prs3$slide_layouts[[6]]
for (i in 1:3) prs3$slides$add_slide(layout)
id_last <- prs3$slides[[3]]$slide_id
prs3$slides$move(prs3$slides[[3]], 1L)
prs3$slides[[1]]$slide_id == id_last  # TRUE
```

## Core properties

```{r}
prs4 <- pptx_presentation()
cp <- prs4$core_properties
cp$title   <- "My Presentation"
cp$author  <- "Jane Doe"
cp$subject <- "rpptx demo"
cp$title
```

## Unit conversion reference

| Function | Units | EMU |
|---|---|---|
| `Emu(n)` | EMU | n |
| `Inches(n)` | inches | n × 914 400 |
| `Cm(n)` | centimetres | n × 360 000 |
| `Pt(n)` | points | n × 12 700 |
| `Mm(n)` | millimetres | n × 36 000 |
| `Centipoints(n)` | centipoints | n × 127 |
