---
title: "Working with Shapes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working with Shapes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup}
library(rpptx)
```

```{r helper, include=FALSE}
blank_slide <- function() {
  prs    <- pptx_presentation()
  layout <- prs$slide_layouts[[6]]  # "Blank"
  list(prs = prs, slide = prs$slides$add_slide(layout))
}
```

## Shape types

Every shape on a slide has a `shape_type` that identifies it:

| Constant | Value | Element |
|---|---|---|
| `MSO_SHAPE_TYPE$AUTO_SHAPE` | 1 | Autoshape (`p:sp`) |
| `MSO_SHAPE_TYPE$CHART` | 3 | Chart (`p:graphicFrame`) |
| `MSO_SHAPE_TYPE$GROUP` | 6 | Group shape (`p:grpSp`) |
| `MSO_SHAPE_TYPE$PICTURE` | 13 | Picture (`p:pic`) |
| `MSO_SHAPE_TYPE$PLACEHOLDER` | 14 | Placeholder (`p:sp`) |
| `MSO_SHAPE_TYPE$TABLE` | 19 | Table (`p:graphicFrame`) |
| `MSO_SHAPE_TYPE$TEXT_BOX` | 17 | Text box (`p:sp`) |
| `MSO_SHAPE_TYPE$LINE` | 9 | Connector (`p:cxnSp`) |

## Common shape properties

All shapes share these properties:

```{r}
r <- blank_slide()
slide <- r$slide
shape <- slide$shapes$add_shape(
  MSO_AUTO_SHAPE_TYPE$RECTANGLE,
  left = Inches(1), top = Inches(1),
  width = Inches(3), height = Inches(2)
)

shape$name         # "Rectangle 1"
shape$shape_id     # integer ID unique within the slide
shape$left         # position in EMU
shape$top
shape$width
shape$height
shape$shape_type   # MSO_SHAPE_TYPE$AUTO_SHAPE
```

Positions and sizes are read/write:

```{r}
shape$left   <- Inches(2)
shape$top    <- Inches(3)
shape$width  <- Inches(4)
shape$height <- Inches(1.5)
```

## Autoshapes

`add_shape()` accepts any `MSO_AUTO_SHAPE_TYPE` constant:

```{r}
r <- blank_slide()
slide <- r$slide

rect  <- slide$shapes$add_shape(MSO_AUTO_SHAPE_TYPE$RECTANGLE,
           Inches(0.5), Inches(0.5), Inches(2), Inches(1.5))
oval  <- slide$shapes$add_shape(MSO_AUTO_SHAPE_TYPE$OVAL,
           Inches(3),   Inches(0.5), Inches(2), Inches(1.5))
arrow <- slide$shapes$add_shape(MSO_AUTO_SHAPE_TYPE$RIGHT_ARROW,
           Inches(5.5), Inches(0.5), Inches(3), Inches(1.5))
```

Some commonly used shape types:

```{r}
MSO_AUTO_SHAPE_TYPE$RECTANGLE
MSO_AUTO_SHAPE_TYPE$ROUNDED_RECTANGLE
MSO_AUTO_SHAPE_TYPE$OVAL
MSO_AUTO_SHAPE_TYPE$DIAMOND
MSO_AUTO_SHAPE_TYPE$TRIANGLE
MSO_AUTO_SHAPE_TYPE$RIGHT_ARROW
MSO_AUTO_SHAPE_TYPE$STAR_5_POINT
MSO_AUTO_SHAPE_TYPE$CLOUD
```

## Text boxes

A text box is a shape with no preset geometry:

```{r}
r <- blank_slide()
slide <- r$slide

tb <- slide$shapes$add_textbox(
  Inches(1), Inches(1), Inches(5), Inches(1)
)
tb$text_frame$text <- "Hello from a text box"
tb$shape_type == MSO_SHAPE_TYPE$TEXT_BOX
```

## Pictures

```{r eval=FALSE}
r <- blank_slide()
slide <- r$slide

pic <- slide$shapes$add_picture(
  "photo.jpg",
  left = Inches(1), top = Inches(1),
  width = Inches(4), height = Inches(3)
)
pic$shape_type == MSO_SHAPE_TYPE$PICTURE
```

## Connectors

Connectors link two points with a line:

```{r}
r <- blank_slide()
slide <- r$slide

conn <- slide$shapes$add_connector(
  MSO_CONNECTOR_TYPE$STRAIGHT,
  begin_x = Inches(1), begin_y = Inches(1),
  end_x   = Inches(5), end_y   = Inches(4)
)
conn$shape_type == MSO_SHAPE_TYPE$LINE
```

Connector types: `STRAIGHT`, `ELBOW`, `CURVE`.

## Fill formatting

Shapes (autoshapes, text boxes) expose a `fill` property:

```{r}
r <- blank_slide()
slide <- r$slide
shape <- slide$shapes$add_shape(
  MSO_AUTO_SHAPE_TYPE$RECTANGLE,
  Inches(1), Inches(1), Inches(3), Inches(2)
)

# Solid fill with RGB colour
shape$fill$solid()
shape$fill$fore_color$rgb <- RGBColor(0x4F, 0x81, 0xBD)

# No fill (transparent)
shape$fill$background()
```

Fill types:

| Method | Description |
|---|---|
| `fill$solid()` | Solid colour (set `fore_color$rgb`) |
| `fill$background()` | No fill / transparent |
| `fill$patterned()` | Pattern fill |
| `fill$gradient()` | Gradient fill |

## Line formatting

```{r}
r <- blank_slide()
slide <- r$slide
shape <- slide$shapes$add_shape(
  MSO_AUTO_SHAPE_TYPE$RECTANGLE,
  Inches(1), Inches(1), Inches(3), Inches(2)
)

shape$line$width                  <- Pt(2)
shape$line$color$rgb              <- RGBColor(0xFF, 0x00, 0x00)
shape$line$dash_style             <- MSO_LINE_DASH_STYLE$DASH
```

## Iterating shapes

`SlideShapes` supports 1-based index access and `to_list()`:

```{r}
r <- blank_slide()
slide <- r$slide
slide$shapes$add_shape(MSO_AUTO_SHAPE_TYPE$RECTANGLE,
  Inches(0.5), Inches(0.5), Inches(2), Inches(1))
slide$shapes$add_textbox(
  Inches(3), Inches(0.5), Inches(3), Inches(1))

length(slide$shapes)   # 2
slide$shapes[[1]]$name

for (shape in slide$shapes$to_list()) {
  cat(shape$name, "—", shape$shape_type, "\n")
}
```

## Group shapes

Multiple shapes can be grouped:

```{r}
r <- blank_slide()
slide <- r$slide
s1 <- slide$shapes$add_shape(MSO_AUTO_SHAPE_TYPE$RECTANGLE,
        Inches(1), Inches(1), Inches(2), Inches(1))
s2 <- slide$shapes$add_shape(MSO_AUTO_SHAPE_TYPE$OVAL,
        Inches(2), Inches(2), Inches(2), Inches(1))

grp <- slide$shapes$add_group_shape(list(s1, s2))
grp$shape_type == MSO_SHAPE_TYPE$GROUP
length(grp$shapes)   # 2
```

## Placeholders

Placeholder shapes (from a slide layout) are accessed via `placeholders`:

```{r}
prs    <- pptx_presentation()
layout <- prs$slide_layouts[[1]]   # "Title Slide" layout
slide  <- prs$slides$add_slide(layout)

# List placeholder indices and types
for (ph in slide$placeholders$to_list()) {
  cat(ph$placeholder_format$idx, "—", ph$placeholder_format$type, "\n")
}
```

Set placeholder text:

```{r}
ph1 <- slide$placeholders[[1]]; ph1$text <- "My Title"
ph2 <- slide$placeholders[[2]]; ph2$text <- "My Subtitle"
```

(`[[n]]` on `SlidePlaceholders` accepts a 1-based position, not the OOXML `idx`.)

## Video (movie shapes)

Embed a video file on a slide with `add_movie()`. The shape shows a poster
frame image until the video plays.

```{r eval=FALSE}
mp4 <- system.file("test_files", "dummy.mp4", package = "rpptx")

r     <- blank_slide()
slide <- r$slide

shape <- slide$shapes$add_movie(
  movie_file = mp4,
  left = Inches(1), top = Inches(1),
  width = Inches(6), height = Inches(4)
)
shape$shape_type  # MSO_SHAPE_TYPE$PICTURE
```

By default the poster frame is a generic loudspeaker icon bundled with the
package. Supply your own image with `poster_frame_image`:

```{r eval=FALSE}
img <- system.file("test_files", "monty-truth.png", package = "rpptx")
slide$shapes$add_movie(mp4,
  Inches(1), Inches(1), Inches(6), Inches(4),
  poster_frame_image = img,
  mime_type = "video/mp4"
)
```

The video relationship uses the OOXML dual-relationship pattern required for
Office 2007/2010 compatibility: one `RT.MEDIA` relationship and one `RT.VIDEO`
relationship both pointing to the same media part. Multiple calls with the same
file reuse a single media part (SHA-1 deduplication).
