---
title: "Working with Tables"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working with Tables}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup}
library(rpptx)
```

```{r helper, include=FALSE}
blank_slide <- function() {
  prs    <- pptx_presentation()
  layout <- prs$slide_layouts[[6]]
  list(prs = prs, slide = prs$slides$add_slide(layout))
}
```

## Creating a table

`add_table()` adds an empty table with a given number of rows and columns:

```{r}
r <- blank_slide()
slide <- r$slide

tbl_shape <- slide$shapes$add_table(
  rows   = 4,
  cols   = 3,
  left   = Inches(1),
  top    = Inches(1),
  width  = Inches(8),
  height = Inches(3)
)
tbl <- tbl_shape$table
class(tbl)   # "Table"
```

## Dimensions

```{r}
tbl$rows     # 4
tbl$columns  # 3
```

## Accessing cells

`cell(row, col)` uses 1-based indices:

```{r}
cell <- tbl$cell(1, 1)
class(cell)   # "_Cell"
```

## Writing cell text

```{r}
cell_11 <- tbl$cell(1, 1); cell_11$text <- "Header A"
cell_12 <- tbl$cell(1, 2); cell_12$text <- "Header B"
cell_13 <- tbl$cell(1, 3); cell_13$text <- "Header C"

cell_21 <- tbl$cell(2, 1); cell_21$text <- "Row 1"
cell_22 <- tbl$cell(2, 2); cell_22$text <- "42"
cell_23 <- tbl$cell(2, 3); cell_23$text <- "3.14"
```

## Filling a table from a data frame

```{r}
r <- blank_slide()
slide <- r$slide

df <- data.frame(
  Name    = c("Alice", "Bob", "Carol"),
  Score   = c(92, 85, 78),
  Grade   = c("A", "B+", "B"),
  stringsAsFactors = FALSE
)

nr   <- nrow(df) + 1L   # +1 for header
nc   <- ncol(df)

tbl_shape <- slide$shapes$add_table(
  nr, nc,
  Inches(1), Inches(1), Inches(8), Inches(3.5)
)
tbl <- tbl_shape$table

# Write column headers
for (j in seq_len(nc)) {
  cell <- tbl$cell(1, j)
  cell$text <- names(df)[j]
}

# Write data rows
for (i in seq_len(nrow(df))) {
  for (j in seq_len(nc)) {
    cell <- tbl$cell(i + 1L, j)
    cell$text <- as.character(df[i, j])
  }
}
```

## Cell fill

```{r}
r <- blank_slide()
slide <- r$slide
tbl_shape <- slide$shapes$add_table(3, 3, Inches(1), Inches(1), Inches(7), Inches(3))
tbl <- tbl_shape$table

# Blue header row
for (j in 1:3) {
  cell <- tbl$cell(1, j)
  cell$fill$solid()
  cell$fill$fore_color$rgb <- RGBColor(0x4F, 0x81, 0xBD)
}

# Alternating row shading
for (i in 2:3) {
  for (j in 1:3) {
    cell <- tbl$cell(i, j)
    if (i %% 2 == 0) {
      cell$fill$solid()
      cell$fill$fore_color$rgb <- RGBColor(0xDB, 0xE5, 0xF1)
    } else {
      cell$fill$background()
    }
  }
}
```

## Cell text frame

Each cell has a `text_frame` for rich text formatting:

```{r}
r <- blank_slide()
slide <- r$slide
tbl_shape <- slide$shapes$add_table(2, 2, Inches(1), Inches(1), Inches(6), Inches(2))
tbl <- tbl_shape$table

# Bold header
cell <- tbl$cell(1, 1)
cell$text <- "Total"
cell$text_frame$paragraphs[[1]]$font$bold <- TRUE

# Right-aligned number
cell2 <- tbl$cell(2, 1)
cell2$text <- "1,234"
cell2$text_frame$paragraphs[[1]]$alignment <- PP_PARAGRAPH_ALIGNMENT$RIGHT
```

## Column widths and row heights

```{r}
r <- blank_slide()
slide <- r$slide
tbl_shape <- slide$shapes$add_table(3, 3, Inches(1), Inches(1), Inches(8), Inches(3))
tbl <- tbl_shape$table

# Set column widths
tbl$columns[[1]]$width <- Inches(3)
tbl$columns[[2]]$width <- Inches(2)
tbl$columns[[3]]$width <- Inches(3)

# Set row heights
tbl$rows[[1]]$height <- Inches(0.6)
tbl$rows[[2]]$height <- Inches(1.0)
tbl$rows[[3]]$height <- Inches(1.0)
```

## Merging cells

`merge()` spans a rectangular block of cells:

```{r}
r <- blank_slide()
slide <- r$slide
tbl_shape <- slide$shapes$add_table(3, 4, Inches(0.5), Inches(1), Inches(9), Inches(3))
tbl <- tbl_shape$table

# Merge the top row across all 4 columns
cell_11 <- tbl$cell(1, 1)
cell_14 <- tbl$cell(1, 4)
cell_11$merge(cell_14)
cell_11$text <- "Merged Header"
```

## Full example

```{r}
prs    <- pptx_presentation()
layout <- prs$slide_layouts[[6]]
slide  <- prs$slides$add_slide(layout)

# Sample data
sales <- data.frame(
  Region  = c("North", "South", "East", "West", "Total"),
  Q1      = c(234, 187, 312, 145, 878),
  Q2      = c(256, 201, 298, 167, 922),
  Q3      = c(289, 223, 341, 189, 1042),
  stringsAsFactors = FALSE
)

nr <- nrow(sales) + 1L
nc <- ncol(sales)

ts <- slide$shapes$add_table(
  nr, nc,
  Inches(0.5), Inches(0.5),
  Inches(9),   Inches(5)
)
tbl <- ts$table

# Headers
header_color <- RGBColor(0x17, 0x37, 0x5E)
for (j in seq_len(nc)) {
  cell <- tbl$cell(1, j)
  cell$text <- names(sales)[j]
  cell$fill$solid()
  cell$fill$fore_color$rgb <- header_color
  tf <- cell$text_frame
  tf$paragraphs[[1]]$font$bold  <- TRUE
  tf$paragraphs[[1]]$font$color$rgb <- RGBColor(0xFF, 0xFF, 0xFF)
}

# Data rows
for (i in seq_len(nrow(sales))) {
  row_color <- if (i %% 2 == 1) RGBColor(0xD6, 0xDC, 0xE4) else RGBColor(0xFF, 0xFF, 0xFF)
  is_total  <- sales$Region[i] == "Total"
  for (j in seq_len(nc)) {
    cell <- tbl$cell(i + 1L, j)
    cell$text <- as.character(sales[i, j])
    cell$fill$solid()
    cell$fill$fore_color$rgb <- if (is_total) header_color else row_color
    if (is_total) {
      cell$text_frame$paragraphs[[1]]$font$bold  <- TRUE
      cell$text_frame$paragraphs[[1]]$font$color$rgb <- RGBColor(0xFF, 0xFF, 0xFF)
    }
    if (j > 1) {
      cell$text_frame$paragraphs[[1]]$alignment <- PP_PARAGRAPH_ALIGNMENT$RIGHT
    }
  }
}

# Column widths
tbl$columns[[1]]$width <- Inches(2)
for (j in 2:nc) tbl$columns[[j]]$width <- Inches(7 / (nc - 1))

tmp <- tempfile(fileext = ".pptx")
prs$save(tmp)
```
